|

### 2020-05-28 paper/Manuscript.Rmd
Integrate AL comments

<!-- I would suggest to put OSCU as last item in the list of affiliations, as it is complementary to the primary institutional affiliations --> 
affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2" 
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "6"
    institution: "University Library Utrecht"
  - id: "7"
    institution: "Utrecht University, Department of Information and Computing Sciences"
  - id: "8"
    institution: "UMC Utrecht"
  - id: "9"
    institution: "Utrecht University, Department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS"

abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to lower the threshold to adoption. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make an entire <!-- Here I would include a word for describing the kind of project. "data science" maybe? --> project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. To facilitate widespread adoption, the
  WORCS principles are implemented in the R-package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

@@

# Introduction

The (social) sciences <!-- Is the "(social)" necessary here? Does it add anything? How about just "science", or alternatively "research" or "academia"? --> are arguably past the tipping point of a paradigm shift
towards open science.
Several immediate causes rallied support for this transition, including
highly publicized cases of scientific fraud [as discussed by
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science should not be seen as a cure (or punishment) for this
crisis. As the late dr. <!-- Is it common to include the title? Maybe rather a description of who he was, his function? --> Jonathan Tennant put it: "Open science is just good science"
[@tennantOpenScienceJust2018].
Open science creates opportunities for researchers to more easily
conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012;
@adolphOpenBehavioralScience2012]. Open science also
promotes inclusivity, because it removes barriers for participation.
Capitalizing on these advances has the potential to accelerate scientific
progress [see also @coyneReplicationInitiativesWill2016].

Many researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
apply open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The Workflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility <!-- here another good point to say something about the scope -->. The workflow (Figure <!--\@ref(fig:workflow)-->) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
The R-package `worcs` offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

@@

WORCS relies on several free, open source software solutions.
<!-- maybe give a short outline of this section here -->

@@

Although there is a slight learning curve to transitioning to DDG, the
investment really pays off. Time saved from painstakingly copy-pasting output
and manually formatting text soon outweighs the investment of switching to a new
program. Moreover, human error in manually copying results is eliminated.
When revisions require major changes to the analyses, all results, figures and
tables can be automatically updated. The flexibility in output formats also 
means that a manuscript can be rendered to presentation format, or even to a
website or blog post. Moreover, reproducibility is guaranteed <!-- that is a strong claim, and probably not true. For example, if the code calls a remote service, that might change its interface or discontinue operation, and the code will not be reproducible anymore - so, suggestion is to weeken this claim a bit... -->, and can be
verified by reviewers and/or readers, because the code is run each time the
document is compiled. In sum, while writing academic papers in a programming environment might seem
counter-intuitive at first, this approach is much more amenable to the needs of
academics than most word processing software. It prevents mistakes, and saves
time. <!-- https://www.scientificpaperofthefuture.org/ could be referenced here -->

@@

The second solution is *version control*:
Maintaining an indelible log of every change to all project files. Version 
control is a near-essential tool for scientific reproducibility, as anyone
learns who has had the experience of accidentally deleting a crucial file, or of
being unable to reproduce analyses because they ran some analyses interactively
instead of using syntax <!-- Also interactive analyses often use syntax, so this is not a good distinction.  "... instead of storing commands in a script file" or something like that would be more precise --> [see also @blischakQuickIntroductionVersion2016].
Many scientists use some form of *implicit* version control; for example, by
renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"),
tracking changes in word processing software, or using cloud hosting services
that retain backups of previous versions.

@@

```{r results="asis", echo = FALSE, eval = TRUE}
get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
  txt <- readLines(filename)
  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
  txt <- gsub("(<!--S: |-->.*$)", "", txt)
  txt <- gsub("^#", "", txt)
  out <- vector("character")
  for(i in seq_along(txt)){
    if(startsWith(txt[i], "#")){
      out <- append(out, c("", txt[i], ""))
    } else {
      out <- append(out, txt[i])
    }
  }
  cat(out, sep = "\n")
}
get_workflow()
```
<!-- Some questions on the workflow, for simplicity I put them here: 
Step 6: "a script documenting the steps" - is that an executable script, or just documentation, or a script with comments? 
Step 7: What about external data sources, actuall? For example, when I use the daily corona stats from Johns Hopkins university directly from their online source. Should they be copied or just a link provided? For FAIR (or at least F) data (with a PID) a link would suffice, but maybe for other sources a copy is advisable. 
Step 14: Is that optional, too? And could it not be done earlier, already in Phase 1?-->

@@

The R-package `worcs` automate most steps of the workflow
for R users.
Several arguments support the choice to implement this workflow in R.
First, R and all of it extensions are free open source software, which
should be the tool of choice for open science.
Second, as of this writing, R is the most advanced programming language in terms
of the implementation of tools required for an open science workflow. <!-- I would omit the second argument. It is highly debatable, and as it stands not more than a claim. Without having looked at the details, I would for example also see Python absolutely capable of doing this. The other four arguments are completely sufficient for motivating the focus on R, so IMHO no reason to open this discussion. --> 
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].

@@

References for software are sometimes provided within the software environment;
for example, in R, package references can be found by
calling `citation("packagename")`. This returns an APA-style reference, and a
BibTeX entry. Software papers are also common, and are sometimes called
"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
offers programmers a way to generate a publication 
based on well-documented software;
in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
@rosenbergTidyLPAPackageEasily2018.
Software citations can also sometimes be found 
in the source code repository (e.g., on 'Git' remote repositories), using the
['Citation File Format'](https://citation-file-format.github.io/).
<!-- just today the Software Heritage Project announced https://www.softwareheritage.org/2020/05/26/citing-software-with-style/ - might also be worth mentioning here? -->

One important impediment to comprehensive citation is the fact that print journals operate with space constraints.
Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
Researchers can overcome this impediment by preparing two versions of the manuscript:
One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
The print version should reference the online version, so interested readers can find the comprehensive reference list.
The WORCS procedure suggests uploading the online version to a preprint server.
This is important because most major preprint servers - at least those hosted <!-- are these "most"? not sure this remark is very accurate --> by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
This means that authors will receive credit for cited work; even if they are cited only in the online version.
Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

@@

As the purpose of the `prepare_data.R` file is to prepare data for sharing in a
format as close to "raw data" as possible, the final line of this file
should typically be either
`open_data()` or `closed_data()`. 
The analysis code, on the other hand, should begin with `load_data()`. <!-- Is this actually limited to one data file, or can it handle multiple ones? --> 
This function loads the real data if it is available on the user's computer, and
otherwise, loads the synthetic data. This will have the effect that third party
users who copy a remote repository with closed data will automatically load the synthetic
dataset, whereas the study authors, who have the real data stored locally, will
automatically load the real data. This function makes it possible for reviewers,
coauthors, and auditors, to write analysis scripts without requiring access to
the original data. They can simply start the script with `load_data()`, and
write their code based on the synthetic data. Then, they can submit their code
to you - by email or as a
["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
When you run their code on your system, `load_data()` will detect the
original data, and use that to run their script. You can then return the results
to the third party.

As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
As explained above, this is recommended, because `.csv` files are human- and
machine readable, and because data typically need to be in a spreadsheet format
for analysis. Many types of data can be represented as a spreadsheet - text
corpora can have one row per document, and EEG- or EKG waveforms can have one
row per measurement occasion. The `prepare_data.R` file should document any
steps required to convert these data into tabular format. If it is data must be
stored in a different file type anyway <!-- something is wrong with the (beginning of) this sentence -->, readers are encouraged to follow the
development of the `repro` package [@repro], which will enable such enhancements.

@@

Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define
preregistration as "creating a permanent record of your study plans before you
look at the data. The plan is stored in a date-stamped, uneditable file in a
secure online archive." Two such archives are well-known in the social sciences:
[AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
However, 'Git' cloud hosting services also conform to these standards. <!-- I am not familiar with these standard, but AFAIK the hosting services make no guarantees regarding long-term availablity of the repositories, in particular not about stable URLs pointing to repositories. They will not disappear suddenly or in the next few years, probably, but if it is important for preregistrations to be available "forever", then they should be made at a service that provides a permanent storage and identifier.  -->
Thus, it is possible to preregister a study simply by pushing a text document
with the study plans to a 'Git' remote repository.
By tagging the commit as a preregistration
(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
it is distinct from all other commits, and easily findable by people and programs.

@@

To preregister a project using 'Git', one can commit
all preregistration documents to the local 'Git' repository, 
and push them 
to a remote 'Git' hosting service.
By [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration" on the remote repository, it can be easily identified as
such and retrieved. This simple but robust approach renders pre-registration through
preregistration services, such as AsPredicted.org or OSF.io, optional.
However, it is easy to take this optional step when the preregistration is
written using dynamic document generation, as 
recommended in WORCS: This file can be formatted as PDF and uploaded as an attachment
to the formal preregistration service.

@@

WORCS was designed to meet the TOP-guidelines, and it does so in a way that
is compatible with other open science principles. Most notably, WORCS is 
amenable to the
FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable. Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well, most notably software [@lamprechtFAIRPrinciplesResearch2019].
WORCS helps users meet the FAIR principles by making all created research objects -
including data and code - accessible in a repository that can be made public. 
The `worcs` package additionally
facilitates meeting the FAIR principles in several ways. First, each `worcs`
project has a YAML file (https://yaml.org/) that lists the research objects in
the project, which will allow web crawlers 
to index `worcs` projects. Second, a codebook is automatically generated for
each data file stored
using `open_data()` or `closed_data()`. We encourage users to elaborate on the
default codebook by adding
variable description and categories, which would - in the future - enable
indexing data by topic area. 
Incidentally, `worcs` itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
hosted on a version controlled public remote repository 'GitHub', has an open
source licence (GPL v3.0), is registered in a community registry (CRAN),
enables the citation of the software (using the `citation("worcs")` command, or
by citing this paper), and followed the CII Best Practices <!-- reference? --> software quality
checklist during development.

@@

A different class of solutions instead focuses on the practical issue of *how*
researchers can meet the requirements of open science.
One notable example is the workflow for
reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
excellent <!-- I am not opposing this statement, but given that the two authors of this work are now co-authors on this very paper, it is a bit awkward to use this qualification... ;) --> workflow strives to ensure strict computational reproducibility for
even the most sophisticated analyses. It has some limitations, however: In
focusing on reproducibility, it only addresses TOP-guidelines 2, 3, and
5. Moreover, in aiming for strict computational reproducibility it ends up being
relatively complex to implement. WORCS builds upon the same general principles
as Peikert and Brandmaier <!-- just noting that their names are frequently repeated. has their approach/tool/... a name that could be used instead? -->, which means the two workflows are compatible. What 
sets WORCS apart is that is an easy and sufficient solution for projects that
can be conducted entirely within R, resulting in a much more lightweight
approach to safeguard computational reproducibility under *most circumstances*.
WORCS also addresses a unique issue not covered by other existing solutions,
namely to provide a workflow most conducive to satisfying the TOP-guidelines and
FAIR principles, while being compatible with existing requirements.

@@

WORCS provides a user-friendly and
lightweight workflow for open, reproducible research, that meets all
TOP-guidelines <!-- was it not with one exception? then maybe just replace the "all" in this sentence by "the" -->. Nevertheless, there are clear directions for future developments.
Firstly, although the workflow is currently implemented only in R, it is 
conceptually relevant for users of other statistical programming languages, most
notably Python [@10.5555/1593511].
We welcome efforts to implement WORCS in other platforms.
Secondly, even when a project is in R, it may have dependencies outside of the
R environment that cannot be managed using `renv`.
It is beyond the scope of the `worcs` package to support the incorporation of
tools outside of R, or to containerize a project so that it can
be identically reinstated on a different system or virtual machine.
Therefore, another
important area of future development is to enable such extensions of the
workflow. To this end, we are collaborating with @repro on the development of
`repro`, an R-package that will facilitate setting up reproducibility tools such
as 'Make' and 'Docker'.
This will make it possible for researchers
to enhance a WORCS project with these strict reproducibility tools,
or vice versa, to initiate a WORCS project within a strictly reproducible
environment generated using `repro`. 
This will combine the strengths of WORCS with those of the workflow proposed
by @peikertReproducibleDataAnalysis2019.

>>>>

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2" 
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "6"
    institution: "University Library Utrecht"
  - id: "7"
    institution: "Utrecht University, Department of Information and Computing Sciences"
  - id: "8"
    institution: "UMC Utrecht"
  - id: "9"
    institution: "Utrecht University, Department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS"

abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to lower the threshold to adoption. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. This workflow is primarily targeted towards
  scholars working in R on a private computer,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data - but it is flexible enough to accommodate many
  other scenarios.
  To facilitate widespread adoption, the
  WORCS principles are implemented in the R-package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

@@

Academia is arguably past the tipping point of a paradigm shift
towards open science.
Several immediate causes rallied support for this transition, including
highly publicized cases of scientific fraud [as discussed by
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science should not be seen as a cure (or punishment) for this
crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
[@tennantOpenScienceJust2018].
Open science creates opportunities for researchers to more easily
conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012;
@adolphOpenBehavioralScience2012]. Open science also
promotes inclusivity, because it removes barriers for participation.
Capitalizing on these advances has the potential to accelerate scientific
progress [see also @coyneReplicationInitiativesWill2016].

Many researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
apply open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The Workflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility for scholars working in R on a private computer,
conducting research that involves academic prose, analysis code,
and (optionally) tabular data - but it is flexible enough to accommodate many
other scenarios. The workflow (Figure \@ref(fig:workflow)) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
The R-package `worcs` offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

```{r workflow, fig.cap="Schematic illustration of the WORCS procedure"}
knitr::include_graphics("workflow_graph/workflow.png")
```

@@

WORCS relies on several free, open source software solutions which we will
discuss before introducing the workflow.

@@

Although there is a slight learning curve to transitioning to DDG, the
investment really pays off. Time saved from painstakingly copy-pasting output
and manually formatting text soon outweighs the investment of switching to a new
program. Moreover, human error in manually copying results is eliminated.
When revisions require major changes to the analyses, all results, figures and
tables can be automatically updated. The flexibility in output formats also 
means that a manuscript can be rendered to presentation format, or even to a
website or blog post. Moreover, the fact that code can be run each time the
document is compiled encourages researchers to work reproducibly,
and allows reviewers and/or readers verify reproducibility simply by re-compiling
the document.
In sum, while writing academic papers in a programming environment might seem
counter-intuitive at first, this approach is much more amenable to the needs of
academics than most word processing software. It prevents mistakes, and saves
time.

@@

The second solution is *version control*:
Maintaining an indelible log of every change to all project files. Version 
control is a near-essential tool for scientific reproducibility, as anyone
learns who has had the experience of accidentally deleting a crucial file, or of
being unable to reproduce analyses because they ran some analyses interactively,
instead of documenting the syntax in a script file [see also @blischakQuickIntroductionVersion2016].
Many scientists use some form of *implicit* version control; for example, by
renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"),
tracking changes in word processing software, or using cloud hosting services
that retain backups of previous versions.

@@

```{r results="asis", echo = FALSE, eval = TRUE}
get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
  txt <- readLines(filename)
  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
  txt <- gsub("(<!--S: |-->.*$)", "", txt)
  txt <- gsub("^#", "", txt)
  out <- vector("character")
  for(i in seq_along(txt)){
    if(startsWith(txt[i], "#")){
      out <- append(out, c("", txt[i], ""))
    } else {
      out <- append(out, txt[i])
    }
  }
  cat(out, sep = "\n")
}
get_workflow()
```

@@

The R-package `worcs` automate most steps of the workflow
for R users.
Several arguments support the choice to implement this workflow in R.
First, R and all of it extensions are free open source software, which
should be the tool of choice for open science.
Second, all tools required for an open science workflow are implemented in R,
most of these tools are directly accessible through the user interface of RStudio,
and some of them are actively developed by the team behind RStudio.
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].

@@

References for software are sometimes provided within the software environment;
for example, in R, package references can be found by
calling `citation("packagename")`. This returns an APA-style reference, and a
BibTeX entry. Software papers are also common, and are sometimes called
"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
offers programmers a way to generate a publication 
based on well-documented software;
in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
@rosenbergTidyLPAPackageEasily2018.
Software citations can also sometimes be found 
in the source code repository (e.g., on 'Git' remote repositories), using the
['Citation File Format'](https://citation-file-format.github.io/). The
[*Software Heritage Project*](https://www.softwareheritage.org/2020/05/26/citing-software-with-style/)
recently published a BibLaTeX software citation style that will further facilitate
accurate citation of software in dynamically generated documents.

One important impediment to comprehensive citation is the fact that print journals operate with space constraints.
Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
Researchers can overcome this impediment by preparing two versions of the manuscript:
One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
The print version should reference the online version, so interested readers can find the comprehensive reference list.
The WORCS procedure suggests uploading the online version to a preprint server.
This is important because most major preprint servers - including [arXiv.org](https://arxiv.org)
and all preprint services hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
This means that authors will receive credit for cited work; even if they are cited only in the online version.
Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

@@

As the purpose of the `prepare_data.R` file is to prepare data for sharing in a
format as close to "raw data" as possible, the final line of this file
should typically be either
`open_data()` or `closed_data()`. 
The analysis code, on the other hand, should begin with `load_data()`.
This function loads the real data if it is available on the user's computer, and
otherwise, loads the synthetic data. This will have the effect that third party
users who copy a remote repository with closed data will automatically load the synthetic
dataset, whereas the study authors, who have the real data stored locally, will
automatically load the real data. This function makes it possible for reviewers,
coauthors, and auditors, to write analysis scripts without requiring access to
the original data. They can simply start the script with `load_data()`, and
write their code based on the synthetic data. Then, they can submit their code
to you - by email or as a
["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
When you run their code on your system, `load_data()` will detect the
original data, and use that to run their script. You can then return the results
to the third party.

Note that these functions can be used to store multiple data files, if necessary.
As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
As explained above, this is recommended, because `.csv` files are human- and
machine readable, and because data typically need to be in a spreadsheet format
for analysis. Many types of data can be represented as a spreadsheet - text
corpora can have one row per document, and EEG- or EKG waveforms can have one
row per measurement occasion. The `prepare_data.R` file should document any
steps required to convert these data into tabular format. 
If this is not an appropriate format, readers are encouraged to follow the
development of the `repro` package [@repro], which will offer greater support for
reproducibly storing and retrieving diverse data formats.

@@

Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define
preregistration as "creating a permanent record of your study plans before you
look at the data. The plan is stored in a date-stamped, uneditable file in a
secure online archive." Two such archives are well-known in the social sciences:
[AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
However, 'Git' cloud hosting services also conform to these standards.
Thus, it is possible to preregister a study simply by committing a preregistration document
to the local 'Git' repository, and pushing it to the remote repository.
Subsequently [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases)
as "Preregistration" on the remote repository renders it distinct from all other commits,
and easily findable by people and programs.
This approach is simple, quick, and robust.
Moreover, it is compatible with formal preregistration through services
such as AsPredicted.org or OSF.io:
If the preregistration is
written using dynamic document generation, as 
recommended in WORCS, this file can be rendered to PDF and uploaded as an attachment
to the formal preregistration service.

@@

@@

WORCS was designed to meet the TOP-guidelines, and it does so in a way that
is compatible with other open science principles. Most notably, WORCS is 
amenable to the
FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable. Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well, most notably software [@lamprechtFAIRPrinciplesResearch2019].

WORCS helps users meet the FAIR principles by making all created research objects -
including data and code - accessible in a repository that can be made public. 
The `worcs` package additionally
facilitates meeting the FAIR principles in several ways. First, each `worcs`
project has a YAML file (https://yaml.org/) that lists the research objects in
the project, which will allow web crawlers 
to index `worcs` projects. Second, a codebook is automatically generated for
each data file stored
using `open_data()` or `closed_data()`. We encourage users to elaborate on the
default codebook by adding
variable description and categories, which would - in the future - enable
indexing data by topic area.

Incidentally, `worcs` itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
hosted on a version controlled public remote repository 'GitHub', has an open
source licence (GPL v3.0), is registered in a community registry (CRAN),
enables the citation of the software (using the `citation("worcs")` command, or
by citing this paper), and followed the [*CII Best Practices*](https://bestpractices.coreinfrastructure.org/en)
software quality checklist during development.

@@

A different class of solutions instead focuses on the practical issue of *how*
researchers can meet the requirements of open science.
One notable example is the workflow for
reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
workflow strives to ensure strict computational reproducibility for
even the most sophisticated analyses. It has some limitations, however: In
focusing on reproducibility, it only addresses TOP-guidelines 2, 3, and
5. Moreover, in aiming for strict computational reproducibility it ends up being
relatively complex to implement. WORCS builds upon the same general principles
as Peikert and Brandmaier,
which means the two workflows are compatible. What 
sets WORCS apart is that is an easy and sufficient solution for projects that
can be conducted entirely within R, resulting in a much more lightweight
approach to safeguard computational reproducibility under *most circumstances*.
WORCS also addresses a unique issue not covered by other existing solutions,
namely to provide a workflow most conducive to satisfying the TOP-guidelines and
FAIR principles, while being compatible with existing requirements.

One initiative that WORCS is fully compatible with, is the [*Scientific Paper of
the Future*](https://www.scientificpaperofthefuture.org/). This organization encourages
geoscientists to document data provenance and availability in public repositories,
document software used,
and document the analysis steps taken to derive the results. All of these goals
could be met using WORCS.

@@

WORCS provides a user-friendly and
lightweight workflow for open, reproducible research, that meets all
TOP-guidelines, except replication.
Nevertheless, there are clear directions for future developments.
Firstly, although the workflow is currently implemented only in R, it is 
conceptually relevant for users of other statistical programming languages, most
notably Python [@10.5555/1593511].
We welcome efforts to implement WORCS in other platforms.
Secondly, even when a project is in R, it may have dependencies outside of the
R environment that cannot be managed using `renv`.
It is beyond the scope of the `worcs` package to support the incorporation of
tools outside of R, or to containerize a project so that it can
be identically reinstated on a different system or virtual machine.
Therefore, another
important area of future development is to enable such extensions of the
workflow. To this end, we are collaborating with @repro on the development of
`repro`, an R-package that will facilitate setting up reproducibility tools such
as 'Make' and 'Docker'.
This will make it possible for researchers
to enhance a WORCS project with these strict reproducibility tools,
or vice versa, to initiate a WORCS project within a strictly reproducible
environment generated using `repro`. 
This will combine the strengths of WORCS with those of the workflow proposed
by @peikertReproducibleDataAnalysis2019.
